---
title: "Final Proposal"
author: "Thanh Mai, Brian Xicon"
output: html_document
date: "2023-11-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description:
The dataset we are using is a list of all active businesses from the City of Los Angeles. The fields include the account number, business name, DBA name, street address, city, zip code, location description, mailing address, mailing city, mailing zip code, primary NAICS description (what the company does), NAICS (code of what the company does), council district, location start date, location end date, and lastly the actual location as coordinates. It has more than 580000 companies in this dataset. We got this dataset from catalog.data.gov
```{r}
businesses <- read.csv("Listing_of_Active_Businesses.csv")
```


# First Data Cleaning Steps:
```{r}
library(tidyverse)

#Converting dates to Date objects
businesses$LOCATION.START.DATE <- as.Date(businesses$LOCATION.START.DATE, format = "%m/%d/%Y")
businesses$LOCATION.END.DATE <- as.Date(businesses$LOCATION.END.DATE, format = "%m/%d/%Y")

#Removing duplicate rows
businesses <- businesses %>% distinct()

head(businesses)
```


# Marginal Summaries:
```{r}
#Some basic numerical marginal summaries
numeric_summaries <- businesses %>%
  summarise(
    StartDateEarliest = min(LOCATION.START.DATE, na.rm = TRUE),
    StartDateLatest = max(LOCATION.START.DATE, na.rm = TRUE),
  )
numeric_summaries

#Some basic categorical marginal summaries
categorical_summaries <- businesses %>%
  filter(!is.na(PRIMARY.NAICS.DESCRIPTION) & PRIMARY.NAICS.DESCRIPTION != "") %>%
  summarise(
    MostCommonCity = names(which.max(table(CITY))),
    MostCommonNAICS = names(which.max(table(PRIMARY.NAICS.DESCRIPTION)))
  )

categorical_summaries
```

